The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2019-06-08 Fsing/FThresholdsPaper/FThresholds.tex
Changes to paper

\documentclass{amsart}
\usepackage{calc,amssymb,amsthm,amsmath,amsfonts,setspace}
\RequirePackage[dvipsnames,usenames]{xcolor}

@@

In summary,  the $F$-pure threshold is an interesting numerical invariant, and is related to many other such measures of singularity across all characteristics.  Unfortunately, it is typically difficult to calculate, even in many concrete cases.  Nevertheless, $F$-pure thresholds have been the focus of intense study over the past 15 years, and the package \emph{FrobeniusThresholds} is centered on calculating and estimating the $F$-pure threshold and other $F$-thresholds, with the function {\tt fpt} at its core.

@@

In the literature, the $F$-pure threshold $\ft^{\idealm}(f)$ is often denoted $\fpt(f)$, for obvious reasons.  However, in this note, we adopt the former notation to avoid any possible confusion with the \emph{Macaulay2} function {\tt fpt} described in Section \ref{sec.FPT}.  Indeed, the command {\tt fpt(f)} need not always output the number $\ft^{\idealm}(f) = \fpt(f)$, though in the case that it cannot, it will provide a reasonably good estimate for this number.

@@

\begin{itemize}
\item[$\circ$] \autoref{sec.Nu} describes the fundamental function {\tt nu}.  As its name suggests, this function is concerned with computing the integers $\nu_I^J(p^e)$ mentioned above.
\item[$\circ$] \autoref{sec.IsFPT} explains the functionality for determining whether a given number is the $F$-pure threshold, or more generally, an \emph{$F$-jumping exponent}.
\item[$\circ$] \autoref{sec.FPT} describes the features of the central function of the package, {\tt fpt}.
\item[$\circ$]  \autoref{sec.FutureDirections} outlines our plans for future improvements to this package.
\end{itemize}
%{\color{red} Move this later}

@@

\section{The {\tt nu} function}
\label{sec.Nu}

We first consider the {\tt nu} function, a fundamental component of the package \emph{FrobeniusThresholds}.   In this section, we adopt the notation established in the introduction.  Namely, we work in a polynomial ring over a finite field of characteristic $p>0$,  $\idealm$ denotes the ideal generated by the variables, and $e$ is a natural number.

If $f$ is a nonzero polynomial contained in the radical of an ideal $J$, then the command {\tt nu(e,f,J)} outputs the integer $\nu_f^J(p^e)$ defined in the introduction, and
%, the maximal integer exponent such that $f^n$ is not contained in $J^{[p^e]}$.
more generally, if a nonzero ideal $I$ is contained in the radical of $J$, then {\tt nu(e,I,J)} outputs $\nu_I^J(p^e)$. Consider the following example.  %, the maximal integer exponent $n$ such that $I^n$ is not contained in $J^{[p^e]}$.

@@

We highlight some degenerate cases:  If $f$ is the zero polynomial, then {\tt nu(e,f,J)} outputs {\tt 0}, and if $f$ is \emph{not} contained in the radical of $J$, then the output is instead {\tt infinity}.  These degenerate cases extend analogously to the command {\tt nu(e,I,J)}.

@@

If the third argument is omitted from the function {\tt nu}, it is assumed to be the maximal ideal $\idealm$.

@@

This implies that when searching for the maximal exponent defining {\tt nu(e,I,J)} for positive $e$, it is safe to start at $p$ times the output of {\tt nu(e-1,I,J)}, and one needn't search too far past this number.  This also suggests that the most efficient way to compute {\tt nu(e,I,J)} is to compute, successively, {\tt nu(s,I,J)}, for each integer $s = 0,\ldots,e$.  This is indeed how the computation is done in most cases.

As pointed out in the introduction, if $f \in \mathfrak{m}$, the values $\nu^{\mathfrak{m}}_f(p^e)$ can be recovered from the $F$-pure threshold of $f$.
This fact is used to speed up computations for certain polynomials whose $F$-pure thresholds can be computed quickly via specialized algorithms, namely diagonal polynomials, binomials, and forms in two variables (see Section \ref{sec.FPT}).  This feature can be disabled by setting the option {\tt UseSpecialAlgorithms} (default value {\tt true}) to {\tt false}.  Consider the following example, where $f$ is a diagonal polynomial.

@@

\subsection{Options for {\tt nu}}

We now describe the options available for the function {\tt nu}.  We focus primarily on the case where the second input is an ideal $I$, since the options work analogously when it is instead a polynomial $f$.

In general, the function {\tt nu} works by searching through a list of integers $n$, and checking containments of the $n$-th power of $I$ in a specified Frobenius power of $J$.
The user can specify how this search is approached through the  option {\tt Search}.  Namely, a binary search is used when {\tt Search} is set to {\tt Binary} (and this is the option's default value), and a linear search is run when it is set to {\tt Linear}.

@@

If the option {\tt ReturnList} is changed from its default value of {\tt false} to {\tt true}, rather than simply outputting $\nu_I^J(p^e)$, {\tt nu} outputs a list of the values $\nu_I^J(p^s)$, for $s=0,\ldots,e$.
Alternatively, the same information can be found by setting the option {\tt Verbose} to {\tt true} (its default value is {\tt false}), to request that each of these values $\nu_I^J(p^s)$ is printed as they are iteratively computed.  This can also serve a way to monitor the progress of the computation.

i8 : nu(5, f, ReturnList => true)

@@

i9 : nu(5, f, Verbose => true)

     nu(1) = 0
     nu(p^1) = 1
     nu(p^2) = 8
     nu(p^3) = 44
     nu(p^4) = 224
     nu(p^5) = 1124
\end{MyVerbatim}
}
\medspace

@@

As described in the introduction, the integer $\nu_I^J(p^e)$ is the maximal integer $n$ such that the $n$-th power of $I$ does not lie in the $p^e$-th Frobenius power of $J$.  However, 
\begin{equation*}
  I^n \subseteq J^{[p^e]} \Longleftrightarrow (I^n)^{[1/p^e]} \subseteq J
\end{equation*}
where $(I^n)^{[1/p^e]}$ denotes the $p^e$-th \emph{Frobenius root} of $I^n$, as defined in \cite{BlickleMustataSmithDiscretenessAndRationalityOfFThresholds}.
The option {\tt ContainmentTest} for {\tt nu} allows the user to choose which of the two types of containment statements appearing above to use toward the calculation of $\nu_I^J(p^e)$.

If {\tt ContainmentTest} is set to {\tt StandardPower}, then {\tt nu(e,I,J)} is computed by testing the left-hand containment above, and when it is set to {\tt FrobeniusRoot}, the right-hand containment is checked.
For efficiency reasons, the default value for {\tt ContainmentTest} is set to  {\tt FrobeniusRoot} if the second argument passed to {\tt nu} is a polynomial, and is set to {\tt StandardPower} if it is an ideal.

@@

The option {\tt ContainmentTest} has a third possible value, {\tt FrobeniusPower}, which may produce a different result than the other two values.
Given an ideal $I$, the first, third, and fourth authors introduced the notion of a (generalized) Frobenius power $I^{[n]}$ of $I$, when $n$ is an arbitrary positive integer, not necessarily of the form $n = p^e$ \cite{hernandez+etal.frobenius_powers}.
When {\tt ContainmentTest} is set to {\tt FrobeniusPower}, rather than  computing $\nu_I^J(p^e)$, the function {\tt nu} computes the maximal integer $n$ for which $I^{[n]}$ is not contained in $J^{[p^e]}$.  This number is denoted $\mu_I^J(p^e)$ in \emph{loc.\ cit.}, and equals $\nu_I^J(p^e)$ when $I$ is replaced with a polynomial, or is a principal ideal.  However, these number need not agree in general, as we see below.

@@

The last option we describe for {\tt nu} is {\tt IsLocal}.
Recall that $\nu_I^\idealm(p^e)$ can be interpreted as the maximum integer $n$ for which $(I^n)^{[1/p^e]}$ is not contained in $\idealm$. 
When computing its value globally, it is the maximum integer $n$ for which $(I^n)^{[1/p^e]} \neq R$.  When 
the value of {\tt IsLocal} is set to {\tt false} (from its default value {\tt true}), then {\tt nu} instead determines this maximal integer $n$, which can also be characterized as the minimal integer $\nu_I^\mathfrak{n}(p^e)$ as $\mathfrak{n}$ varies among all maximal ideals of $R$.

@@

%%Below is the original version of the section describing the nu function

%\section{The $F$-pure threshold limit and {\tt nu}}
%\label{sec.Nu}
%
%Suppose $R = k[x_1, \dots, x_n]$ is a polynomial ring over a finite field and $f \in R$.    If $f^a \notin \langle x_1^{p^e}, \dots, x_n^{p^e} \rangle = \fram^{[p^e]}$, then the pair $(R, f^{a/(p^e -1)})$ is $F$-pure at the origin, whereas if $f^a \in \fram^{[p^e]}$, then the pair $(R, f^{a/p^e})$ is not $F$-pure at the origin.  By finding the $a \in \bZ$ where that transition happens, and then limiting over $e$, one can compute the $F$-pure threshold.  See \cite{MustataTakagiWatanabeFThresholdsAndBernsteinSato}.
%
%We set $\nu(e, f)$ to be largest $a \in \bZ$ such that $f^a \notin \fram^{[p^e]}$ (note in the literature, it is frequently denoted by $\nu_f(p^e)$.  It immediately follows that
%\[
%\fpt(f) = \lim_{e \to \infty} {\nu(e, f) \over p^e}.
%\]
%These numbers $\nu(e, f)$ can be computed using the command {\tt nu}, as in the example below.
%\medskip
%{\small
%\setstretch{.67}
%\begin{MyVerbatim}
%i1 : loadPackage "FThresholds"
%
%o1 = FThresholds
%
%o1 : Package
%
%i2 : R = ZZ/7[x,y]
%
%o2 = R
%
%o2 : PolynomialRing
%
%i3 : f = y^2 - x^3
%
%        3    2
%o3 = - x  + y
%
%o3 : R
%
%i4 : nu(1, f)
%
%o4 = 5
%
%o4 : QQ
%
%i5 : nu(2, f)
%
%o5 = 40
%
%o5 : QQ
%
%i6 : nu(3, f)
%
%o6 = 285
%
%o6 : QQ
%\end{MyVerbatim}
%}
%\medskip
%The numbers $5/7, 40/49, 285/343$ then approximate the $F$-pure threshold, which in this case happens to equal $5/6$.  By default, if the function $f$ has a special form (is a diagonal polynomial or a binomial), then the function does not check whether $f^a$ is in  $\langle x_1^{p^e}, \dots, x_n^{p^e} \rangle$, but instead uses general formulas for $\nu$s, see \cite{HernandezFPureThresholdOfBinomial,HernandezFInvariantsOfDiagonalHyp}.
%
%In fact, one can make the following more general definition.  If $I$ and $J$ are two ideals (with $I \subseteq \sqrt{J}$), one can define
%\[
%\nu(e, I, J) = \max \{ n \in \bZ \; |\; I^n \not\subseteq J^{[p^e]} \}.
%\]
%This is frequently denoted by $\nu_I^J(p^e)$ in the literature.  In the case that $I = \langle f \rangle$ and $J = \fram$, this just recovers the $\nu$'s defined above; the limit of which is the $F$-pure threshold.  For more general $I$ and $J$, the limit
%\[
%\lim_{e \to \infty} {\nu(e, I, J) \over p^e}
%\]
%is called the \emph{$F$-threshold of $I$ with respect to $J$}.  To compute these values we use a binary search, unless the user specifies that we should use a linear search instead {\tt Search=>Linear}.
%
%There is one more option that deserves mentioning, {\tt ContainmentTest}.  There are two ways to check whether $f^a$ (or $I^a$) is contained in $J^{[p^e]}$.  In general it is true that
%\[
%I^a \subseteq J^{[p^e]} \text{ $\Leftrightarrow$ } (I^a)^{[1/p^e]} \subseteq J
%\]
%where $\bullet^{[1/p^e]}$ is defined as in the {\tt TestIdeals} package.  Which of these two strategies to use is controlled by the option {\tt ContainmentTest => StandardPower} and {\tt ContainmentTest => FrobeniusRoot} respectively.
%For principal $I$, the command on the right is generally faster (and so {\tt FrobeniusRoot} is the default strategy for computing $\nu$).  For non-principal $I$, frequently the command on the left is quicker, and so {\tt StandardPower} is default strategy.
%
%Finally, the option {\tt ContainmentTest => FrobeniusPower}, then ${\tt \nu(e, I, A)}$ computes the largest integer $n$ such that
%\[
%I^{[p^n]} \subseteq J^{[p^e]}.
%\]
%See \cite{HernandezTeixeiraWittFrobeniusPowers} for more discussion of this.

\section{{\tt isFPT}, {\tt compareFPT} and {\tt isFJumpingExponent}}
\label{sec.IsFPT}

@@

Consider a  $\mathbb{Q}$-Gorenstein ring $R$ of characteristic $p>0$, whose index not divisible by $p$.
Given a parameter $t \in \mathbb{Q}$ and an element $f$ of $R$, the command {\tt isFPT(t,f)} outputs {\tt true} if $t$ is the $F$-pure threshold of $f$, and {\tt false} otherwise.
Moreover, the function {\tt compareFPT} provides further information about the relationship between $t$ and the $F$-pure threshold of $f$; namely,
{\tt compareFPT(t, f)} returns {\tt-1}, {\tt 0}, or {\tt 1} when $t$ is less than, equal to, or greater than, respectively, the $F$-pure threshold of $f$.  Setting the option {\tt IsLocal} to {\tt true} tells the function to consider the $F$-pure threshold at the origin.  The following example illustrates this functionality.

@@

The general method applied calls upon functionality from the \emph{TestIdeals} package.  The functions {\tt testIdeal} and {\tt FPureModule} therein are used
to compute the test ideals of $f$ with respect to the parameter $t$, and that with respect to $t - \varepsilon$, for $\varepsilon>0$ very small. For instance, as noted in the introduction, when $R$ is a polynomial ring, $t$ is the $F$-pure threshold of $f$ if and only if $\tau(R, f^t)$ is contained in the homogeneous maximal ideal, but
$\tau(R, f^{t-\varepsilon})$ is not.
Note that for $R$ strongly $F$-regular and Gorenstein, {\tt FPureModule(t, f)} precisely computes $\tau(R, f^{t-\varepsilon})$.

@@

\section{The {\tt fpt} function}
\label{sec.FPT}

The core function in the package \emph{FrobeniusThresholds} is called {\tt fpt}.  Throughout this section, let $f$ be a polynomial with coefficients in a finite field of characteristic $p$. When passed the polynomial $f$, the function {\tt fpt} attempts to find the exact value for the $F$-pure threshold of $f$ at the origin, and returns that value, if possible.  Otherwise, it returns lower and upper bounds for the $F$-pure threshold, as demonstrated below.

@@

\subsection{The option {\tt UseSpecialAlgorithms}}

The {\tt fpt} function has an option called {\tt UseSpecialAlgorithms}, which, when set to {\tt true} (its default value), tells {\tt fpt} to first check whether $f$ is a diagonal polynomial, a binomial, a form in two variables, or a product of factors in simple normal crossing, in that order.

@@

The above noted algorithm for computing the $F$-pure threshold of a binary form $f$ requires factoring $f$ into linear forms, and {\tt fpt} can sometimes hang when attempting that factorization, when the factorization occurs in a Galois field of excessively large order.
This is a situation when the user will want to set the option \texttt{UseSpecialAlgorithms} to \texttt{false}.
However, when a factorization is already known, instead of passing $f$ to {\tt fpt}, the user can pass a list containing all the pairwise coprime linear factors of $f$ to {\tt fpt}, and a list containing their respective multiplicities.

@@

If {\tt UseSpecialAlgorithms} is set to {\tt true} and $f$ does not fall into any of the aforementioned cases, then the function {\tt fpt} next calls {\tt isSimpleNormalCrossing(f)} to check whether the polynomial $f$ is a product of factors that are in simple normal crossing, in which case the $F$-pure threshold can easily be computed.
For more detail on the function {\tt isSimpleNormalCrossing} and on computing the $F$-pure threshold of a polynomial with simple normal crossings, please see Section \ref{subsec.SNC}.

@@

In the remainder of this section, we explain how the function  {\tt fpt} proceeds when no special algorithm is available, or when {\tt UseSpecialAlgorithms} is set to {\tt false}, and describe the roles of the options {\tt DepthOfSearch} and {\tt Attempts}, whose values are both nonnegative integers.

In this case, the {\tt fpt} function either computes, and outputs, the exact value of the $F$-pure threshold of $f$, or otherwise recursively computes a finite sequence of lower and upper bounds for this value, and outputs the last of these, which will be the tightest among all computed.  The value of the option {\tt DepthOfSearch} determines the precision of the initial set of bounds, and the option {\tt Attempts}, in conjunction with some various subroutines, determines whether, and how, to produce new, tighter bounds from the previous ones.

More specifically, let $e$ denote the value of the option {\tt DepthOfSearch}, which conservatively defaults to {\tt 1}.  The {\tt fpt} function first computes $\nu=\nu_f(p^e)$, which agrees with the output of {\tt nu(e,f)}.  It is well known that the $F$-pure threshold of $f$ is greater than $\nu/p^e$ and at most $(\nu+1)/p^e$, and applying  \cite[Proposition 4.2]{HernandezFPurityOfHypersurfaces} to this lower bound tells us that the $F$-pure threshold of $f$ must be at least $\nu/(p^e-1)$.  In summary, we know that the $F$-pure threshold of $f$ must lie in the closed interval
%
\begin{equation}
\label{estimating-interval: e}
\tag{$\dagger$}
\left[ \frac{\nu}{p^e-1}, \frac{\nu+1}{p^e} \right].
\end{equation}

With these estimates in hand, the subroutine {\tt guessFPT} is called to make some ``educated guesses" in an attempt to identify the $F$-pure threshold within this interval, or at least narrow down this interval to produce improved estimates.  The number of ``guesses" is controlled by the option {\tt Attempts}, which conservatively defaults to {\tt 3}.  If {\tt Attempts} is set to {\tt 0}, then {\tt guessFPT} is bypassed. If  {\tt Attempts} is set to at least {\tt 1}, then a first check is run to verify whether the right-hand endpoint $(\nu+1)/p^e$ of the above interval \eqref{estimating-interval: e} is the $F$-pure threshold.  We illustrate this below.

@@

If  {\tt Attempts} is set to at least {\tt 2} and the right-hand endpoint $(\nu+1)/p^e$ of the interval \eqref{estimating-interval: e} is not the $F$-pure threshold, then a second check is run to verify whether the left-hand endpoint $\nu/(p^e-1)$ of this interval is the $F$-pure threshold.

@@

If neither endpoint is the $F$-pure threshold and {\tt Attempts} is set to more than {\tt 2}, then  additional checks are performed at certain numbers within the interval.
First, a number in the interval is selected, according to criteria specified by the value of the option {\tt GuessStrategy}; we refer the reader for the documentation of this option for more details.
Then the function {\tt compareFPT} is used to test that number. If that ``guess'' is correct, its value is returned; otherwise, the data returned by {\tt compareFPT} is used to narrow down the interval, and this process is repeated as many times as specified by {\tt Attempts}.

@@

The option {\tt Bounds} allows the user to specify known lower and upper bounds for the $F$-pure threshold of $f$, in order to speed up computations or to refine previously obtained estimates.

@@

If {\tt guessFPT} is unsuccessful and {\tt FinalAttempt} is set to {\tt true}, the {\tt fpt} function proceeds to use the convexity of the $F$-signature function and a secant line argument to attempt to narrow down the interval bounding the $F$-pure threshold.
If successful, the new lower bound may coincide with the upper bound, in which case we can conclude that it is the desired $F$-pure threshold.
If this is not the case, an $F$-regularity check is done at the new lower bound, to verify if it is the $F$-pure threshold.

@@

The computations performed when {\tt FinalAttempt} is set to {\tt true} are often slow, and often fail to improve the estimate, and for this reason, this option should be used sparingly. It is often more effective to increase the values of {\tt Attempts} or {\tt DepthOfSearch}, instead.

@@

Finally, we point out that one can set the option {\tt IsLocal} from its default value of {\tt true} to {\tt false}, to compute the $F$-pure threshold globally. In other words, it computes the minimum of the $F$-pure threshold at all maximal ideals.

>>>>

\documentclass{amsart}
\usepackage{calc,amssymb,amsthm,amsmath,amsfonts,setspace}
\RequirePackage[dvipsnames,usenames]{xcolor}
\usepackage{textgreek}
\usepackage[utf8x]{inputenc}
\DeclareUnicodeCharacter{957}{\textnu}

@@

In summary,  the $F$-pure threshold is an interesting numerical invariant, and is related to many other such measures of singularity across all characteristics.  Unfortunately, it is typically difficult to calculate, even in many concrete cases.  Nevertheless, $F$-pure thresholds have been the focus of intense study over the past 15 years, and the package \emph{FrobeniusThresholds} is centered on calculating and estimating the $F$-pure threshold and other $F$-thresholds, with the function \texttt{fpt} at its core.

@@

In the literature, the $F$-pure threshold $\ft^{\idealm}(f)$ is often denoted $\fpt(f)$, for obvious reasons.  However, in this note, we adopt the former notation to avoid any possible confusion with the \emph{Macaulay2} function \texttt{fpt} described in Section \ref{sec.FPT}.  Indeed, the command \texttt{fpt(f)} need not always output the number $\ft^{\idealm}(f) = \fpt(f)$, though in the case that it cannot, it will provide a reasonably good estimate for this number.

@@

\begin{itemize}
\item[$\circ$] \autoref{sec.Nu} describes the fundamental function \texttt{nu}.  As its name suggests, this function is concerned with computing the integers $\nu_I^J(p^e)$ mentioned above.
\item[$\circ$] \autoref{sec.IsFPT} explains the functionality for determining whether a given number is the $F$-pure threshold, or more generally, an \emph{$F$-jumping exponent}.
\item[$\circ$] \autoref{sec.FPT} describes the features of the central function of the package, \texttt{fpt}.
\item[$\circ$]  \autoref{sec.FutureDirections} outlines our plans for future improvements to this package.
\end{itemize}
%{\color{red} Move this later}

@@

\section{The \texttt{nu} function}
\label{sec.Nu}

We first consider the \texttt{nu} function, a fundamental component of the package \emph{FrobeniusThresholds}.   In this section, we adopt the notation established in the introduction.  Namely, we work in a polynomial ring over a finite field of characteristic $p>0$,  $\idealm$ denotes the ideal generated by the variables, and $e$ is a natural number.

If $f$ is a nonzero polynomial contained in the radical of an ideal $J$, then the command \texttt{nu(e,f,J)} outputs the integer $\nu_f^J(p^e)$ defined in the introduction, and
%, the maximal integer exponent such that $f^n$ is not contained in $J^{[p^e]}$.
more generally, if a nonzero ideal $I$ is contained in the radical of $J$, then \texttt{nu(e,I,J)} outputs $\nu_I^J(p^e)$. Consider the following example.  %, the maximal integer exponent $n$ such that $I^n$ is not contained in $J^{[p^e]}$.

@@

We highlight some degenerate cases:  If $f$ is the zero polynomial, then \texttt{nu(e,f,J)} outputs \texttt{0}, and if $f$ is \emph{not} contained in the radical of $J$, then the output is instead \texttt{infinity}.  These degenerate cases extend analogously to the command \texttt{nu(e,I,J)}.

@@

If the third argument is omitted from the function \texttt{nu}, it is assumed to be the maximal ideal $\idealm$.

@@

This implies that when searching for the maximal exponent defining \texttt{nu(e,I,J)} for positive $e$, it is safe to start at $p$ times the output of \texttt{nu(e-1,I,J)}, and one needn't search too far past this number.  This also suggests that the most efficient way to compute \texttt{nu(e,I,J)} is to compute, successively, \texttt{nu(s,I,J)}, for each integer $s = 0,\ldots,e$.  This is indeed how the computation is done in most cases.

As pointed out in the introduction, if $f \in \mathfrak{m}$, the values $\nu^{\mathfrak{m}}_f(p^e)$ can be recovered from the $F$-pure threshold of $f$.
This fact is used to speed up computations for certain polynomials whose $F$-pure thresholds can be computed quickly via specialized algorithms, namely diagonal polynomials, binomials, and forms in two variables (see Section \ref{sec.FPT}).  This feature can be disabled by setting the option \texttt{UseSpecialAlgorithms} (default value \texttt{true}) to \texttt{false}.  Consider the following example, where $f$ is a diagonal polynomial.

@@

\subsection{Options for \texttt{nu}}

We now describe the options available for the function \texttt{nu}.  We focus primarily on the case where the second input is an ideal $I$, since the options work analogously when it is instead a polynomial $f$.

In general, the function \texttt{nu} works by searching through a list of integers $n$, and checking containments of the $n$-th power of $I$ in a specified Frobenius power of $J$.
The user can specify how this search is approached through the  option \texttt{Search}.  Namely, a binary search is used when \texttt{Search} is set to \texttt{Binary} (and this is the option's default value), and a linear search is run when it is set to \texttt{Linear}.

@@

If the option \texttt{ReturnList} is changed from its default value of \texttt{false} to \texttt{true}, rather than simply outputting $\nu_I^J(p^e)$, \texttt{nu} outputs a list of the values $\nu_I^J(p^s)$, for $s=0,\ldots,e$.
Alternatively, the same information can be found by setting the option \texttt{Verbose} to \texttt{true} (its default value is \texttt{false}), to request that each of these values $\nu_I^J(p^s)$ is printed as they are iteratively computed.  This can also serve a way to monitor the progress of the computation.

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}
i8 : nu(5, f, ReturnList => true)

@@

i9 : nu(5, f, Verbose => true)
nuInternal: using comparison test FrobeniusRoot
ν(1) = 0
ν(p^1) = 1
ν(p^2) = 8
ν(p^3) = 44
ν(p^4) = 224
ν(p^5) = 1124

o9 = 1124
\end{MyVerbatim}
}
\medspace

@@

As described in the introduction, the integer $\nu_I^J(p^e)$ is the maximal integer $n$ such that the $n$-th power of $I$ does not lie in the $p^e$-th Frobenius power of $J$.  However, 
\begin{equation*}
  I^n \subseteq J^{[p^e]} \Longleftrightarrow (I^n)^{[1/p^e]} \subseteq J
\end{equation*}
where $(I^n)^{[1/p^e]}$ denotes the $p^e$-th \emph{Frobenius root} of $I^n$, as defined in \cite{BlickleMustataSmithDiscretenessAndRationalityOfFThresholds}.
The option \texttt{ContainmentTest} for \texttt{nu} allows the user to choose which of the two types of containment statements appearing above to use toward the calculation of $\nu_I^J(p^e)$.

If \texttt{ContainmentTest} is set to \texttt{StandardPower}, then \texttt{nu(e,I,J)} is computed by testing the left-hand containment above, and when it is set to \texttt{FrobeniusRoot}, the right-hand containment is checked.
For efficiency reasons, the default value for \texttt{ContainmentTest} is set to  \texttt{FrobeniusRoot} if the second argument passed to \texttt{nu} is a polynomial, and is set to \texttt{StandardPower} if it is an ideal.

@@

The option \texttt{ContainmentTest} has a third possible value, \texttt{FrobeniusPower}, which may produce a different result than the other two values.
Given an ideal $I$, the first, third, and fourth authors introduced the notion of a (generalized) Frobenius power $I^{[n]}$ of $I$, when $n$ is an arbitrary positive integer, not necessarily of the form $n = p^e$ \cite{hernandez+etal.frobenius_powers}.
When \texttt{ContainmentTest} is set to \texttt{FrobeniusPower}, rather than  computing $\nu_I^J(p^e)$, the function \texttt{nu} computes the maximal integer $n$ for which $I^{[n]}$ is not contained in $J^{[p^e]}$.  This number is denoted $\mu_I^J(p^e)$ in \emph{loc.\ cit.}, and equals $\nu_I^J(p^e)$ when $I$ is replaced with a polynomial, or is a principal ideal.  However, these number need not agree in general, as we see below.

@@

The last option we describe for \texttt{nu} is \texttt{IsLocal}.
Recall that $\nu_I^\idealm(p^e)$ can be interpreted as the maximum integer $n$ for which $(I^n)^{[1/p^e]}$ is not contained in $\idealm$. 
When computing its value globally, it is the maximum integer $n$ for which $(I^n)^{[1/p^e]} \neq R$.  When 
the value of \texttt{IsLocal} is set to \texttt{false} (from its default value \texttt{true}), then \texttt{nu} instead determines this maximal integer $n$, which can also be characterized as the minimal integer $\nu_I^\mathfrak{n}(p^e)$ as $\mathfrak{n}$ varies among all maximal ideals of $R$.

@@

\section{\texttt{isFPT}, \texttt{compareFPT} and \texttt{isFJumpingExponent}}
\label{sec.IsFPT}

@@

Consider a  $\mathbb{Q}$-Gorenstein ring $R$ of characteristic $p>0$, whose index not divisible by $p$.
Given a parameter $t \in \mathbb{Q}$ and an element $f$ of $R$, the command \texttt{isFPT(t,f)} outputs \texttt{true} if $t$ is the $F$-pure threshold of $f$, and \texttt{false} otherwise.
Moreover, the function \texttt{compareFPT} provides further information about the relationship between $t$ and the $F$-pure threshold of $f$; namely,
\texttt{compareFPT(t, f)} returns {\tt-1}, \texttt{0}, or \texttt{1} when $t$ is less than, equal to, or greater than, respectively, the $F$-pure threshold of $f$.  Setting the option \texttt{IsLocal} to \texttt{true} tells the function to consider the $F$-pure threshold at the origin.  The following example illustrates this functionality.

@@

The general method applied calls upon functionality from the \emph{TestIdeals} package.  The functions \texttt{testIdeal} and \texttt{FPureModule} therein are used
to compute the test ideals of $f$ with respect to the parameter $t$, and that with respect to $t - \varepsilon$, for $\varepsilon>0$ very small. For instance, as noted in the introduction, when $R$ is a polynomial ring, $t$ is the $F$-pure threshold of $f$ if and only if $\tau(R, f^t)$ is contained in the homogeneous maximal ideal, but
$\tau(R, f^{t-\varepsilon})$ is not.
Note that for $R$ strongly $F$-regular and Gorenstein, \texttt{FPureModule(t, f)} precisely computes $\tau(R, f^{t-\varepsilon})$.

@@

\section{The \texttt{fpt} function}
\label{sec.FPT}

The core function in the package \emph{FrobeniusThresholds} is called \texttt{fpt}.  Throughout this section, let $f$ be a polynomial with coefficients in a finite field of characteristic $p$. When passed the polynomial $f$, the function \texttt{fpt} attempts to find the exact value for the $F$-pure threshold of $f$ at the origin, and returns that value, if possible.  Otherwise, it returns lower and upper bounds for the $F$-pure threshold, as demonstrated below.

@@

\subsection{The option \texttt{UseSpecialAlgorithms}}

The \texttt{fpt} function has an option called \texttt{UseSpecialAlgorithms}, which, when set to \texttt{true} (its default value), tells \texttt{fpt} to first check whether $f$ is a diagonal polynomial, a binomial, a form in two variables, or a product of factors in simple normal crossing, in that order.

@@

The above noted algorithm for computing the $F$-pure threshold of a binary form $f$ requires factoring $f$ into linear forms, and \texttt{fpt} can sometimes hang when attempting that factorization, when the factorization occurs in a Galois field of excessively large order.
This is a situation when the user will want to set the option \texttt{UseSpecialAlgorithms} to \texttt{false}.
However, when a factorization is already known, instead of passing $f$ to \texttt{fpt}, the user can pass a list containing all the pairwise coprime linear factors of $f$ to \texttt{fpt}, and a list containing their respective multiplicities.

@@

If \texttt{UseSpecialAlgorithms} is set to \texttt{true} and $f$ does not fall into any of the aforementioned cases, then the function \texttt{fpt} next calls \texttt{isSimpleNormalCrossing(f)} to check whether the polynomial $f$ is a product of factors that are in simple normal crossing, in which case the $F$-pure threshold can easily be computed.
For more detail on the function \texttt{isSimpleNormalCrossing} and on computing the $F$-pure threshold of a polynomial with simple normal crossings, please see Section \ref{subsec.SNC}.

@@

In the remainder of this section, we explain how the function  \texttt{fpt} proceeds when no special algorithm is available, or when \texttt{UseSpecialAlgorithms} is set to \texttt{false}, and describe the roles of the options \texttt{DepthOfSearch} and \texttt{Attempts}, whose values are both nonnegative integers.

In this case, the \texttt{fpt} function either computes, and outputs, the exact value of the $F$-pure threshold of $f$, or otherwise recursively computes a finite sequence of lower and upper bounds for this value, and outputs the last of these, which will be the tightest among all computed.  The value of the option \texttt{DepthOfSearch} determines the precision of the initial set of bounds, and the option \texttt{Attempts}, in conjunction with some various subroutines, determines whether, and how, to produce new, tighter bounds from the previous ones.

More specifically, let $e$ denote the value of the option \texttt{DepthOfSearch}, which conservatively defaults to \texttt{1}.  The \texttt{fpt} function first computes $\nu=\nu_f(p^e)$, which agrees with the output of \texttt{nu(e,f)}.  It is well known that the $F$-pure threshold of $f$ is greater than $\nu/p^e$ and at most $(\nu+1)/p^e$, and applying  \cite[Proposition 4.2]{HernandezFPurityOfHypersurfaces} to this lower bound tells us that the $F$-pure threshold of $f$ must be at least $\nu/(p^e-1)$.  In summary, we know that the $F$-pure threshold of $f$ must lie in the closed interval
%
\begin{equation}
\label{estimating-interval: e}
\tag{$\dagger$}
\left[ \frac{\nu}{p^e-1}, \frac{\nu+1}{p^e} \right].
\end{equation}

With these estimates in hand, the subroutine \texttt{guessFPT} is called to make some ``educated guesses" in an attempt to identify the $F$-pure threshold within this interval, or at least narrow down this interval to produce improved estimates.  The number of ``guesses" is controlled by the option \texttt{Attempts}, which conservatively defaults to \texttt{3}.  If \texttt{Attempts} is set to \texttt{0}, then \texttt{guessFPT} is bypassed. If  \texttt{Attempts} is set to at least \texttt{1}, then a first check is run to verify whether the right-hand endpoint $(\nu+1)/p^e$ of the above interval \eqref{estimating-interval: e} is the $F$-pure threshold.  We illustrate this below.

@@

If  \texttt{Attempts} is set to at least \texttt{2} and the right-hand endpoint $(\nu+1)/p^e$ of the interval \eqref{estimating-interval: e} is not the $F$-pure threshold, then a second check is run to verify whether the left-hand endpoint $\nu/(p^e-1)$ of this interval is the $F$-pure threshold.

@@

If neither endpoint is the $F$-pure threshold and \texttt{Attempts} is set to more than \texttt{2}, then  additional checks are performed at certain numbers within the interval.
First, a number in the interval is selected, according to criteria specified by the value of the option \texttt{GuessStrategy}; we refer the reader for the documentation of this option for more details.
Then the function \texttt{compareFPT} is used to test that number. If that ``guess'' is correct, its value is returned; otherwise, the data returned by \texttt{compareFPT} is used to narrow down the interval, and this process is repeated as many times as specified by \texttt{Attempts}.

@@

The option \texttt{Bounds} allows the user to specify known lower and upper bounds for the $F$-pure threshold of $f$, in order to speed up computations or to refine previously obtained estimates.

@@

If \texttt{guessFPT} is unsuccessful and \texttt{FinalAttempt} is set to \texttt{true}, the \texttt{fpt} function proceeds to use the convexity of the $F$-signature function and a secant line argument to attempt to narrow down the interval bounding the $F$-pure threshold.
If successful, the new lower bound may coincide with the upper bound, in which case we can conclude that it is the desired $F$-pure threshold.
If this is not the case, an $F$-regularity check is done at the new lower bound, to verify if it is the $F$-pure threshold.

@@

The computations performed when \texttt{FinalAttempt} is set to \texttt{true} are often slow, and often fail to improve the estimate, and for this reason, this option should be used sparingly. It is often more effective to increase the values of \texttt{Attempts} or \texttt{DepthOfSearch}, instead.

@@

Finally, we point out that one can set the option \texttt{IsLocal} from its default value of \texttt{true} to \texttt{false}, to compute the $F$-pure threshold globally. In other words, it computes the minimum of the $F$-pure threshold at all maximal ideals.
